Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic

### 2017-06-30 manuscript/manuscript.tex
expanded Sec. 2

Pulsar timing arrays (PTAs) \sv{blah blah blah}

@@

In this paper, we present a technique for improving the accuracy of the optimal statistic 
by marginalizing the optimal statistic over the individual red noise parameters 
using the posterior distributions from a full Bayesian analysis. 
We demonstrate that this hybrid approach produces a value and uncertainty 
of the optimal statistic that is in better agreement to the $\Agw$ 
produced from a full Bayesian analysis.

@@

In order to compare the noise-marginalized optimal statistic to the 
fixed-noise optimal statistic, we simulated a stochastic background 
in a PTA with \sv{fill in simulation details...} 
The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
To compute the fixed-noise optimal statistic, the red noise parameters are 
fixed to the maximum likelihood values found using \sv{Kristina fill in here...}

@@

\begin{figure}[ht]
	\includegraphics[width=\columnwidth]{plots/os_dataset50.pdf}
	\caption{Optimal statistic and SNR for a simulated dataset 
			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
			We compare the values found using a fixed-noise analysis (blue, dashed line) to the 
			mean optimal statistic and mean SNR found by marginalizing over 
			the individual pulsars' red noise parameters (orange, solid line). 
			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			For this particular realization of the stochastic background, the fixed-noise analysis 
			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$.}
	\label{fig:os_dataset_sample}
\end{figure}
\begin{figure}[ht]
	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15.pdf}
	\caption{Optimal statistic and SNR for \sv{insert number} simulated datasets 
			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
			We compare the values found using a fixed-noise analysis (blue) to the 
			mean optimal statistic and mean SNR found by marginalizing over 
			the individual pulsars' red noise parameters. The dashed vertical line 
			indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			The value of $\hat{A}^2$ from the fixed-noise analysis, 
			averaged over realizations of the stochastic background, is $\hat{A}^2 = 7.9 \times 10^{-30}$, 
			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.2 \times 10^{-29}$.}
	\label{fig:os_datasetstats}
\end{figure}

@@

The optimal statistic

\citep{thk+2016}

@@

Another way to test the spatial correlations is through sky scrambles \sv{blah blah blah...} 
\citet{cs2016} \citet{tlb+2017}

The optimal statistic provides a much faster way to perform sky scrambles compared to 
a Bayesian analysis since it only requires altering the overlap reduction function. 
We performed sky scrambles by \sv{explain here...} 
The scrambled sky positions were generated \sv{Steve, explain here...}

>>>>

Pulsar timing arrays (PTAs) can be used to detect low-frequency gravitational waves (GWs) \sv{blah blah blah}

@@

In this paper, we present a technique for improving the accuracy of the optimal statistic 
by marginalizing the optimal statistic over the individual red noise parameters 
using the posterior distributions from a full Bayesian analysis. 
This hybrid approach produces a more accurate 
estimate of \Agw\ and its uncertainty.

@@

In order to compare the noise-marginalized optimal statistic to the 
fixed-noise optimal statistic, we injected a GW stochastic background with 
$\Agw = 5 \times 10^{-15}$ into simulated PTA data. 
We constructed simulated data sets using the same procedure as in \citet{tlb+2017}, 
which consisted of 10 MSPs with \sv{Steve fill in here...}

The optimal statistic was computed using the \sv{fill in...} in PAL2 \sv{add citation}. 
We incorporate the individual pulsar noise parameters as follows. 
We perform individual Bayesian noise analyses on all of the pulsars, and 
fix the white noise to the maximum likelihood value. 
To compute the fixed-noise optimal statistic, 
the red noise parameters for each pulsar are set to the mode of 
the joint posterior distribution for $A_\mathrm{red}$,$\gamma_\mathrm{red}$, 
found by performing a Bayesian common red noise analysis. 
For the noise-marginalized optimal statistic, we draw from 
the common red noise posteriors in order to average the optimal statistic over 
the individual pulsars' red noise.

@@

\begin{figure}[ht]
	\includegraphics[width=\columnwidth]{plots/os_dataset50.pdf}
	\caption{Optimal statistic and SNR for a simulated dataset 
			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
			We compare the values found using a fixed-noise analysis (blue, dashed line) to the 
			mean optimal statistic and mean SNR found by marginalizing over 
			the individual pulsars' red noise parameters (orange, solid line). 
			The dotted vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			For this particular realization of the stochastic background, the fixed-noise analysis 
			gives $\hat{A}^2 = 1.9 \times 10^{-30}$ with $\mathrm{SNR} = 0.60$, 
			while the noise-marginalized analysis gives $\hat{A}^2 = 1.9\times10^{-29}$ with $\mathrm{SNR} = 1.1$.}
	\label{fig:os_dataset_sample}
\end{figure}

\sv{compare to Bayesian analysis for a particular data set}

\begin{figure}[ht]
	\includegraphics[width=\columnwidth]{plots/os_datasetstats_5e-15.pdf}
	\caption{Optimal statistic and SNR for \sv{insert number} simulated datasets 
			containing an injected GW background with $\Agw = 5\times10^{-15}$. 
			We compare the values found using a fixed-noise analysis (blue) to the 
			mean optimal statistic and mean SNR found by marginalizing over 
			the individual pulsars' red noise parameters. The dashed vertical line 
			indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			The value of $\hat{A}^2$ from the fixed-noise analysis, 
			averaged over realizations of the stochastic background, is $\hat{A}^2 = 7.9 \times 10^{-30}$, 
			while the noise-marginalized analysis gives a mean value of $\hat{A}^2 = 2.2 \times 10^{-29}$.}
	\label{fig:os_datasetstats}
\end{figure}

@@

The optimal statistic is particularly well-suited to compare multiple spatial correlation relations 
because using a different spatial correlation only requires changing the overlap reduction function 
in $\tilde{{\bf{S}}}_{IJ}$. 
\citet{thk+2016} demonstrated how the optimal statistic can be altered to fit for 
multiple spatial correlations at once in order to mitigate common noise sources such as 
clock error and ephemeris error. 
Here we compute the optimal statistic with monopole and dipole spatial correlations 
for the same simulated data sets as in the previous section in order to determine 
how well we can distinguish a GW background from a monopole or dipole signal. 
For a monopole signal, the overlap reduction function becomes simply
\[ \Gamma(\theta_{IJ}) = 1 \,\]
while for a dipole signal, the overlap reduction function becomes
\[ \Gamma(\theta_{IJ}) = \cos\theta_{IJ} \,. \]

Figure~\ref{fig:} shows the mean value of $\hat{A}^2$ and the mean SNR 
computed assuming monopole, dipole, and Hellings-Downs spatial correlations 
for \sv{insert number} simulated data sets.

@@

Another way to test the spatial correlations is through ``sky scrambles,'' 
where the overlap reduction function is altered in order to simulate changing the pulsars' positions \citep{cs2016}. 
\citet{tlb+2017} showed how sky scrambles affect the Bayes' factor for simulated data sets. 
Here we perform a similar analysis using frequentist methods.

The scrambled sky positions were generated \sv{Steve, explain here...}
